\section*{Додаток А. Список публікацій здобувача за темою дисертації}

\begin{enumerate}
	\item Tymchenko B., Neural network methods for planar image analysis in automated screening systems // Applied aspects information technology. 2021. Том 4, № 1. С. 71-79.  (Index Copernicus) 
	
	\textit{https://aait.opu.ua/?fetch=articles\&with=info\&id=73}
	
	\item Tymchenko B., Khvedchenya E., Marchenko P., Spodarets D. Classification of skin lesions using multi-task deep neural networks // Herald of advanced information technology. 2020. Том 3, № 3. С. 136–148.  (Index Copernicus)
	
	\textit{https://hait.opu.ua/?fetch=articles\&with=info\&id=62}
	
	\item Tymchenko B., Marchenko P., Spodarets D. Segmentation of cloud organization patterns from satellite images using deep neural networks // Herald of advanced information technology. 2020. Том 3, № 1. С. 352-361 (Index Copernicus)
	
	\textit{https://hait.opu.ua/\%EF\%BF\%BD\%EF\%BF\%BD\%EF\%BF\%BD\%EF\%BF\%BD\%EF\%BF\%BD\%EF\%BF\%BD\%EF\%BF\%BD\%EF\%BF\%BD\%C3\%A0/?fetch=articles\&with=info\&id=44}
	
	\item Tymchenko B., Marchenko P., Spodarets D. Deep learning approach to diabetic retinopathy detection // Proceedings of the 9th International Conference on Pattern Recognition Applications and Methods, ICPRAM 2020, Valletta, Malta, February 22-24, 2020 / Ed. by Maria De Marsico, Gabriella Sanniti di Baja, Ana L. N. Fred. –– SCITEPRESS, 2020. C. 501–509. (Scopus)
	
	\textit{https://www.scitepress.org/Papers/2020/89708/}
	
	\item Tymchenko B., Hramatik A., Tulchiy H. Classifying mixed patterns of proteins in microscopic images with deep neural networks // Herald of advanced information technology. 2019. Том 2, № 1. С. 29-36 (Index Copernicus)
	
	\textit{https://hait.opu.ua/?fetch=articles\&with=info\&id=18}
	
	\item Tymchenko B., Antoshchuk S. Race from pixels: Evolving neural network controller for vision-based car driving // XVIII International Conference on Data Science and Intelligent Analysis of Information / Springer. 2018. С. 20–29. (Index Copernicus) 
	
	\textit{https://www.springerprofessional.de/en/race-from-pixels-evolving-neural-network-controller-for-vision-b/16003024}
	
	\item Tymchenko B, Hramatik A., Tulchyi H., Antoshchuk S. Making money: Evolving neural network for stock prediction // VI українсько-нiмецька конференцiя «Iнформатика. Культура. Технiка». 2018. С. 34-35
	
	\item Tymchenko B., Antoshchuk S. Evolution strategy for policy search in robotics // Сучаснi Iнформацiйнi Технологiї / Одеський Нацiональний Полiтехнiчний Унiверситет. 2018. 
	
	\textit{http://dspace.opu.ua/jspui/handle/123456789/8056}
	
	\item Tymchenko B., Halchonkov O. Online lane detection algorithm for line scan camera // Сучаснi Iнформацiйнi Технологiї / Одеський Нацiональний Полiтехнiчний Унiверситет. 2017. C. 86-89 
	
	\textit{http://dspace.opu.ua/jspui/handle/123456789/3351}
	
	\item Tymchenko B. Global position system sensor model for robotics simulator // Працi Одеського полiтехнiчного унiверситету. 2017. № 3. С. 88–93. 
	
	\textit{http://dspace.opu.ua/jspui/handle/123456789/7939}
	
	\item Tymchenko B., Samodelok V., Putilina D., Galchonkov O. The robust control system for skid elimination in dynamic road environments // Електротехнiчнi та комп’ютернi системи. 2016. № 23. С. 107–112.
	http://dspace.opu.ua/jspui/handle/123456789/1176
	
	\item Komleva N., Cherneha K., Tymchenko B., Komlevoy O. Intellectual approach application for pulmonary diagnosis 2016 IEEE First International Conference on Data Stream Mining Processing (DSMP). 2016. C. 48–52.
	
	\textit{https://ieeexplore.ieee.org/document/7583505/}
	
	\item Cherneha K., Tymchenko B., Komleva N. Decision support system for automated medical diagnostics // Електротехнiчнi та комп’ютернi системи. 2016. № 23. С. 65–72. (Index Copernicus)
	
	\textit{http://dspace.opu.ua/xmlui/handle/123456789/1140}
	
\end{enumerate}